OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-03-05 paper/paper.Rmd
More numbers

```{r ref_db}
open_fit_mcc <- dat %>% filter(method == 'open', strategy == 'database', tool == 'mothur') %>% pull(mcc_median) %>% median()
opticlust_mcc <- dat %>% filter(method == 'de_novo', tool == 'mothur') %>% pull(mcc_median) %>% median()
open_fit_db_vs_clust <- rel_diff(opticlust_mcc, open_fit_mcc)

open_fit_gg_mcc <- dat %>% filter(method == 'open', strategy == 'database', tool == 'mothur', ref == 'gg') %>% pull(mcc_median) %>% median()
open_vsearch_mcc <- dat %>% filter(method == 'open', strategy == 'database', tool == 'vsearch', ref == 'gg') %>% pull(mcc_median) %>% median()
open_fit_db_vs_vsearch <- rel_diff(open_fit_gg_mcc, open_vsearch_mcc)
```

@@

In open reference mode, fitting the datasets to reference OTUs with OptiFit
produced OTUs of similar quality (`r open_fit_db_vs_clust`% difference in median
MCC) as clustering the datasets _de novo_ with OptiClust  across all datasets
and reference databases.
OptiFit produced higher quality OTUs than VSEARCH when open reference clustering
against the Greengenes database, with median MCC scores of `r open_fit_gg_mcc`
and `r open_vsearch_mcc` (respectively).
However, open reference VSEARCH and OptiClust respectively ran **X%** and **Y%**
faster than OptiFit in open reference mode.
_De novo_ clustering with OptiClust produced **X%** higher quality OTUs than
VSEARCH, but performed **Y%** slower than VSEARCH.

>>>>

```{r ref_db}
open_fit_mcc <- dat %>% filter(method == 'open', strategy == 'database', tool == 'mothur') %>% pull(mcc_median) %>% median()
opticlust_mcc <- dat %>% filter(method == 'de_novo', tool == 'mothur') %>% pull(mcc_median) %>% median()
mcc_open_fit_db_vs_clust <- rel_diff(opticlust_mcc, open_fit_mcc)

open_fit_gg_mcc <- dat %>% filter(method == 'open', strategy == 'database', tool == 'mothur', ref == 'gg') %>% pull(mcc_median) %>% median()
open_vsearch_mcc <- dat %>% filter(method == 'open', strategy == 'database', tool == 'vsearch', ref == 'gg') %>% pull(mcc_median) %>% median()
mcc_open_fit_db_vs_vsearch <- rel_diff(open_fit_gg_mcc, open_vsearch_mcc)

open_vsearch_sec <- dat %>% filter(method == 'open', strategy == 'database', tool == 'vsearch', ref == 'gg') %>% pull(sec_median) %>% median()
open_fit_sec <- dat %>% filter(method == 'open', strategy == 'database', tool == 'mothur') %>% pull(sec_median) %>% median()
opticlust_sec <- dat %>% filter(method == 'de_novo', tool == 'mothur') %>% pull(sec_median) %>% median()
sec_vsearch_vs_open_fit_db <- rel_diff(open_vsearch_sec, open_fit_sec)
sec_opticlust_vs_open_fit_db <- rel_diff(open_fit_sec, opticlust_sec)
```

```{r de_novo}
dn_vsearch_mcc <- dat %>% filter(strategy == 'de_novo', tool == 'vsearch') %>% pull(mcc_median) %>% median()
dn_vsearch_sec <- dat %>% filter(strategy == 'de_novo', tool == 'vsearch') %>% pull(sec_median) %>% median()
mcc_opticlust_vs_vsearch <- rel_diff(opticlust_mcc, dn_vsearch_mcc)
sec_opticlust_vs_vsearch <- rel_diff(opticlust_sec, dn_vsearch_sec)
```

@@

In open reference mode, fitting the datasets to reference OTUs with OptiFit
produced OTUs of similar quality (`r mcc_open_fit_db_vs_clust`% difference in median
MCC) as clustering the datasets _de novo_ with OptiClust  across all datasets
and reference databases.
OptiFit produced higher quality OTUs than VSEARCH when open reference clustering
against the Greengenes database, with median MCC scores of `r open_fit_gg_mcc`
and `r open_vsearch_mcc` (respectively).
However, OptiClust and VSEARCH each ran `r sec_opticlust_vs_open_fit_db`% and 
`r sec_vsearch_vs_open_fit_db`% faster than OptiFit in open reference mode.
_De novo_ clustering with OptiClust produced `r mcc_opticlust_vs_vsearch`%
higher quality OTUs than VSEARCH, but performed `r sec_opticlust_vs_vsearch`%
slower than VSEARCH.
